\documentclass[12pt]{article}
\usepackage{epsfig}
\usepackage{natbib}
\usepackage{multirow}
\usepackage{amsmath}
\bibliographystyle{mybst}
\input boldmath.tex
\bigpage
\newtheorem{theorem}{Theorem}
\begin{document}
\title{Outline 2}
\date{February 14, 2019} \maketitle
\subsection{Step1: }

To calculate the max purity function, first,  
fit the linear mixed model for the outcome $\b{y}$ and time $\bX$, with baseline covariates $\bx$:  
$$\by = \bX(\bbeta +\bb +\bGamma(\balpha'\bx)) + \bepsilon.$$We can define the covariate matrix of $\bX$ as $\bz$. The $\bz$ contains both fixed effects and random effects. 
$$\bz = \beta + \bb + \bGamma x$$
We can also write the above equation in the matrix version:
$$Y = \left[\begin{array}
{rrr}
1 & t_1 & t_1^2 \\
1 & t_2 & t_2^2 \\
1 & t_3 & t_3^2 \\
1 & t_4 & t_4^2
\end{array}\right]
\left[\begin{pmatrix}
\beta_0 \\
\beta_1 \\
\beta_2
\end{pmatrix} +  
\begin{pmatrix}
b_0 \\
b_1 \\
b_2
\end{pmatrix} + 
\begin{pmatrix}
\gamma_0 \\
\gamma_1 \\
\gamma_2
\end{pmatrix} x \right] + \bepsilon$$
where 

* $\bbeta$ is the vector of covariates for fixed effects of $\bX$.

* $\bb$ is the vector of random effects 

* $\bGamma \bx$ is the vector of fixed effects of the baseline covariates.

* $x$ is the combination of the input baseline covariates. $x = \balpha'\bx$ and $\bx$ is the covariates vector.

\subsection{Step2:}

Then we need to estimate the distribution of $\b{z}$ for drug group and placebo, separately.
$$\begin{aligned}
f(z) &= \int_x f(z,x) dx \\
 & = \int_x f(z|x) g(x) dx
\end{aligned}$$
where,

* the conditional distribution 
$f(z|x) \sim MVN( \bbeta + \bGamma \bx, \bD)$

* $g(x)$ is the distribution of the covariates combination. For example, if the covariates combination only contains "sex", which is binary, then the intergal becomes summation.

* $\bD$ is the covariates matrix of random effects $\bb$. 

\\[1cm]
~\\
\noindent
We could then fit the $f(\cdot)$ for drug group and placebo group separately, i.e. $f_1(z)$ and $f_2(z)$. 

\subsection{Step3:}

Define the purity for a combination of covariates x as 
$$f_1(z|x) = |2\pi D|^{-\frac{1}{2}} exp(-\frac{1}{2} ( \bz - \hat{\bbeta}_1 - \hat{\bGamma}_1 x)^{T} D^{-1} (\bz - \hat{\bbeta}_1  - \hat{\bGamma}_1 x) )$$
$$f_2(z|x) = |2\pi D|^{-\frac{1}{2}} exp(-\frac{1}{2} ( \bz - \hat{\bbeta}_1 - \hat{\bGamma}_1 x)^{T} D^{-1} (\bz - \hat{\bbeta}_1  - \hat{\bGamma}_1 x) )$$

$$\int_x (\frac{\pi_1 f_1(z|x) - \pi_2 f_2(z|x)}{\pi_1 f_1(z|x) + \pi_2 f_2(z|x)}})^2 g(x) dx $$
which can be also approximated as the summation form: 
$$\sum_{x} (\frac{\pi_1 f_1(z|x) - \pi_2 f_2(z|x)}{\pi_1 f_1(z|x) + \pi_2 f_2(z|x)}})^2 g(x)$$

~\\
\noindent
This is one purity value based on one combination of baseline covariates. Define the purity as $P_x$

~\\

\noindent
We would like to find the $\balpha$ that max the $P_x$

\bibliography{all}
\end{document}
