---
title: "Simulation"
date: 2019-04-19
output: pdf_document
---


```{R include = FALSE}
library(png)
library(grid)
true_generation = function(alpha, p, n, ni, tt, X, 
                           beta_drg, gamma_drg, bi_sigma,sigma_drg,
                           beta_pbo, gamma_pbo, bi_sigma2,sigma_pbo){
  # alpha
  set.seed(123)
  alpha = as.matrix(alpha,p,1)
  dat_drg = c()
  for(i in 1:n){
    drg_temp = NULL
    drg_temp$subj = rep(paste('drg',i,sep=''),ni)
    drg_temp$trt = rep('drg',ni)
    baseline = as.matrix(rnorm(p,0,1),p,1)
    x1 = baseline[1]; x2 = baseline[2]; x3 =  baseline[3]
    w = rep(t(alpha) %*% baseline,ni)
    drg_temp$x1 = rep(x1,ni); drg_temp$x2 = rep(x2,ni); drg_temp$x3 = rep(x3,ni) 
    drg_temp$w = w
    drg_temp$tt = tt
    bi = mvrnorm(1, c(0,0,0), bi_sigma)
    yi = X%*%(beta_drg+bi+gamma_drg*w[1]) + sigma_drg*rnorm(ni,0,1)
    drg_temp$y = yi
    dat_drg = rbind(dat_drg, as.data.frame(drg_temp))
  }
  
  dat_pbo = c()
  for(i in 1:n){
    pbo_temp = NULL
    pbo_temp$subj = rep(paste('pbo',i,sep=''),ni)
    pbo_temp$trt = rep('pbo',ni)
    baseline = as.matrix(rnorm(p),p,1)
    x1 = baseline[1]; x2 = baseline[2]; x3 =  baseline[3]
    w = rep(t(alpha) %*% baseline,ni)
    pbo_temp$x1 = rep(x1,ni); pbo_temp$x2 = rep(x2,ni); pbo_temp$x3 = rep(x3,ni) 
    pbo_temp$w = w
    pbo_temp$tt = tt
    bi = mvrnorm(1, c(0,0,0), (bi_sigma2))
    yi = X%*%(beta_pbo+bi+gamma_pbo*w[1]) + sigma_pbo*rnorm(ni,0,1)
    pbo_temp$y = yi
    dat_pbo = rbind(dat_pbo, as.data.frame(pbo_temp))
  }
  print('True data generated')
  return(list(dat_drg = dat_drg, dat_pbo = dat_pbo))
}
```


```{R include = FALSE}
library(lme4)
library(MASS)
library(mixtools)
# set parameters:
# generate beta randomly
beta_drg = as.matrix(c(0,-1,-3),3,1)
beta_pbo = as.matrix(c(0,3,1),3,1)

# generate gamma randomly
gamma_drg=matrix(c(0,-2,1),3,1)
gamma_pbo=matrix(c(0,1,-2),3,1)

# bi
set.seed(21)
eign1 = diag(c(1,0.4,0.5))
eign2 = matrix(runif(9,0,1),3,3)
a = eign2 %*% eign1 %*% solve(eign2)
bi_sigma = t(a) %*% (a)
eigen(bi_sigma)

# bi
set.seed(7)
eign1 = diag(c(1,0.5,2))
eign2 = matrix(runif(9,0,2),3,3)
bi_sigma2 = eign2 %*%  t(eign2) /10
eigen(bi_sigma2)

# sigma
sigma_drg = 1
sigma_pbo = 1

tt = as.matrix(0:6) # pt = the order of time points
ni = length(tt) # 7
X = cbind(matrix(1, length(tt), 1), tt, tt^2)

n = 50 # number of subjects in each group  
p = 3 # number of baseline covariates
ni = 7 # number of time points

alpha = c(1,1,1)

data = true_generation(alpha, p, n, ni, tt, X, 
                      beta_drg, gamma_drg, bi_sigma,sigma_drg,
                      beta_pbo, gamma_pbo, bi_sigma2,sigma_pbo)
dat_drg = data$dat_drg
dat_pbo = data$dat_pbo
dat = rbind(dat_drg, dat_pbo)

```


# Highligths

* Try 3 dimensions of the baseline covariates simulation##

* There are several different way to set the true $\alpha$, with penalty or without penalty. If it is with penalty, df = 1, the results similar 

* If df = 3, it is hard to find the max purity and the max purity maybe not at the true $\alpha$ value ($purity = f(\alpha)$)



# Scenario 1. When $\alpha$ value is randomly chosen

##### Step 1: Specify true $\alpha$, where $\alpha$ is a vector with length 3. Here I made the true value $\alpha$:

```{R}
alpha = c(1,1,1)

```

#####  Step 2: Specify $\beta, \Gamma, D$ for drug group (drg) and placebo group (pbo)

The outcome follows the formula:
$$\text{the outcome is } Y: \mathbf{Y} = \mathbf{X}(\mathbf{\beta} +\mathbf{b} +\mathbf{\Gamma(\alpha'x))} + \mathbf{\epsilon}.$$

where 

* The baselines come from the same distributions $\mathbf{x} = [x_1, x_2, x_3]$, the baseline covariate $x_1$, $x_2$, $x_3$ iid $\sim N(0,1)$

* A combination of baseline covariate $w$: $w = \alpha^T  [x_1,x_2, x_3]$, which is $\alpha_1*x_1 + \alpha_2*x2 + \alpha_3*x3$

* $X$: the independent covariates, $X = [1, t, t^2]$, where $t = 0,1,2,...,6$ 


The $\beta$: 

* $\beta_{drg} = \left[\begin{array}
{r}
0 \\
-1 \\
-3 
\end{array}\right]$, $\beta_{pbo} = \left[\begin{array}
{r}
0 \\
3 \\
1 
\end{array}\right]$

The $\Gamma$: 

* $\Gamma_{drg} = \left[\begin{array}
{r}
0 \\
-2 \\
1 
\end{array}\right]$,
$\Gamma_{pbo} = \left[\begin{array}
{c}
0 \\
1 \\
-2 
\end{array}\right]$

The $\epsilon$:

* The $\epsilon_{drg}$, $\epsilon_{pbo}$ $\sim N(0,1)$

The $bi$:

* The random effect covariance matrix $D_{drg}$:

```{R}
set.seed(21)
eign1 = diag(c(1,0.4,0.5))
eign2 = matrix(runif(9,0,1),3,3)
a = eign2 %*% eign1 %*% solve(eign2)
bi_sigma = t(a) %*% (a) # make sure it is positive defined
round(bi_sigma,3)
```

* The random effect covariance matrix $D_{pbo}$:

```{R}
set.seed(7)
eign1 = diag(c(1,0.5,2))
eign2 = matrix(runif(9,0,2),3,3)
bi_sigma2 = eign2 %*%  t(eign2) /10 # make sure it is positive defined
round(bi_sigma2,3)
```


##### Step 3: Simulate $n$ observations (e.g. n = 100/treatment)

For drug group, the baseline covariates are generated from N(0,1):
```{R}
p = 3
baseline = as.matrix(rnorm(p,0,1),p,1)
x1 = baseline[1]; x2 = baseline[2]; x3 = baseline[3]
```

Then $w$ is calculated as $\alpha'x$

```{R}
alpha = c(1, 1, 1)
w = t(alpha) %*% baseline
```

The random effect is generated from MVN, whose mean = 0 and sigma equals to the covariance matrix $D$

```{R}
bi = mvrnorm(1, c(0,0,0), bi_sigma)
```

Then the outcome is calculated
```{R}
yi = X%*%(beta_drg+bi+gamma_drg*w[1]) + sigma_drg*rnorm(ni,0,1)
```

The placebo group follows the same generation process. 

The whole code for true data generation:

```{R}
true_generation = function(alpha, p, n, ni, tt, X, 
                           beta_drg, gamma_drg, bi_sigma,sigma_drg,
                           beta_pbo, gamma_pbo, bi_sigma2,sigma_pbo){
  # alpha
  set.seed(123)
  alpha = as.matrix(alpha,p,1)
  dat_drg = c()
  for(i in 1:n){
    drg_temp = NULL
    drg_temp$subj = rep(paste('drg',i,sep=''),ni)
    drg_temp$trt = rep('drg',ni)
    baseline = as.matrix(rnorm(p,0,1),p,1)
    x1 = baseline[1]; x2 = baseline[2]; x3 =  baseline[3]
    w = rep(t(alpha) %*% baseline,ni)
    drg_temp$x1 = rep(x1,ni); drg_temp$x2 = rep(x2,ni); drg_temp$x3 = rep(x3,ni) 
    drg_temp$w = w
    drg_temp$tt = tt
    bi = mvrnorm(1, c(0,0,0), bi_sigma)
    yi = X%*%(beta_drg+bi+gamma_drg*w[1]) + sigma_drg*rnorm(ni,0,1)
    drg_temp$y = yi
    dat_drg = rbind(dat_drg, as.data.frame(drg_temp))
  }
  
  dat_pbo = c()
  for(i in 1:n){
    pbo_temp = NULL
    pbo_temp$subj = rep(paste('pbo',i,sep=''),ni)
    pbo_temp$trt = rep('pbo',ni)
    baseline = as.matrix(rnorm(p),p,1)
    x1 = baseline[1]; x2 = baseline[2]; x3 =  baseline[3]
    w = rep(t(alpha) %*% baseline,ni)
    pbo_temp$x1 = rep(x1,ni); pbo_temp$x2 = rep(x2,ni); pbo_temp$x3 = rep(x3,ni) 
    pbo_temp$w = w
    pbo_temp$tt = tt
    bi = mvrnorm(1, c(0,0,0), (bi_sigma2))
    yi = X%*%(beta_pbo+bi+gamma_pbo*w[1]) + sigma_pbo*rnorm(ni,0,1)
    pbo_temp$y = yi
    dat_pbo = rbind(dat_pbo, as.data.frame(pbo_temp))
  }
  print('True data generated')
  return(list(dat_drg = dat_drg, dat_pbo = dat_pbo))
}
```


##### Step 4: Monte carlo simulation

Since we would like to calculate the integral of 
$$\int\frac{[f_1(z_i|w_i) - f_2(z_i|w_i)]^2}{f_1(z_i|w_i) + f_2(z_i|w_i)} dz_i$$
where we assume that $f_1(z_i|w_i)$ and $f_2(z_i|w_i)$ are two MVN for drug group and placebo group separately. 

To calculate this, we could firstly generate a large dataset (10000) from a 2 by 2 standard multivariate normal distribution.

```{R}
Xstart = mvrnorm(10000, c(0,0), diag(c(1,1)))
```

Then transform those points to MVN of drug and MVN of placebo separately.

* standard MVN: $X_0 \sim MVN(\mu_0, \sigma_0)$

* MVN_drg: $X_1 \sim MVN(\mu_{drg}, \sigma_{drg})$, $\sigma_{drg} = H\Lambda H^T$, which is the eigenvalue decomposition of $\sigma_{drg}$. $H$ is the matrix of eigen vectors and $\Lambda$ is the  matrix whose diagnoal values are $\sigma_{drg}$'s eignevalues.

* Transformation: $X_1 = H \Lambda^{\frac{1}{2}} X + \mu_{drg}$

And then we can get the large data points sampled from the two MVN distributions. The purity can be calculated. 

```{R}
### monta
monta_carlo_pdf = function(Xstart, mu1, D1, mu2, D2){
 
  # transformation
  d1 = eigen(D1)
  d1 = d1$vectors %*% diag(sqrt(d1$values))
  
  d2 = eigen(D2)
  d2 = d2$vectors %*% diag(sqrt(d2$values))
  
  points1 = Xstart %*% t(d1) 
  
  points1[,1] = points1[,1] + mu1[1]; points1[,2] = points1[,2] + mu1[2]  
  
  points2 = Xstart %*% t(d2) 
  points2[,1] = points2[,1] + mu2[1]; points2[,2] = points2[,2] + mu1[2]  
  
  f1 = dmvnorm(points1,  mu1, D1)
  f2 = dmvnorm(points2,  mu2, D2)
  
  f3 = (f1 + f2) # deal with the 0s, f1 + f2 cannot be 0
  f3 = ifelse(f3 == 0, 1e-4,f3)
  
  purity = c((f1 - f2)^2 / f3) 
  return(purity)
}
```


##### Step 5: Fit model using $\alpha s$ and compute purity

The functions

```{R}
purity_calculation = function(dat, beta1, beta2, gamma1, gamma2, D1, D2, p=2){
  unique_dat = unique(dat[,c('subj','w')])
  purity = c()
  Mu1 = c(); Mu2 = c()
  for(i in 1:dim(unique_dat)[1]){
    #if(i %% 10 ==0) print(i)
    mu1 = beta1 + gamma1 * unique_dat$w[i] # mu1 = beta_drg + gamma_drg * unique_dat$w[i]
    mu2 = beta2 + gamma2 * unique_dat$w[i] # mu2 = beta_pbo + gamma_pbo * unique_dat$w[i]
    Mu1 = rbind(Mu1, mu1); Mu2 = rbind(Mu2, mu2)
    res = mean(monta_carlo_pdf(Xstart, mu1, D1, mu2, D2))
    purity = c(purity, res)
  }
  return(list(purity = purity, Mu1 = Mu1, Mu2 = Mu2))
}
purity_function = function(A, varname = '', times = '', 
                           trt = '',
                           trtlevel = '',
                           subj = '',
                           outcome = '',
                           start = 0, data = dat){
  p = length(A)
  alpha_est = matrix(A,p,1)
  dat_est = data
  if(sum(varname == '') == length(varname)){
    w_est = dat[(start):(start + p -1),] %*% alpha_est
  }
  if(sum(varname != '') == length(varname)){
    w_est = as.matrix(dat_est[,varname]) %*% alpha_est
  }
  dat_est$w = w_est
  if(times != ''){
    dat_est$tt = dat_est[,times]
  }
  if(subj != ''){
    dat_est$subj = dat_est[,subj]
  }
  if(trt != ''){
    dat_est$trt = dat_est[,trt]
    if(sum(trtlevel == '')==0){
      dat_est[dat_est$trt == trtlevel[1],]$trt = 'pbo'
      dat_est[dat_est$trt == trtlevel[2],]$trt = 'drg'
    }
  }
  dat_est$outcome = dat_est[,outcome]
  dat_pbo_est = dat_est[dat_est$trt == 'pbo', ]
  dat_drg_est = dat_est[dat_est$trt == 'drg', ]
  
  fit_drg_est = lmer(outcome ~ tt + I(tt^2) + w + w * tt +
                       w * I(tt^2) + (tt+I(tt^2)|subj),
                     data = dat_drg_est, REML = FALSE)
  fit_drg_est
  fit_pbo_est = lmer(outcome ~ tt + I(tt^2) + w + w * tt +
                       w * I(tt^2) + (tt+I(tt^2)|subj),
                     data = dat_pbo_est, REML = FALSE)
  fit_pbo_est
  
  beta1 = as.matrix(fixef(fit_drg_est))[2:3]
  gamma1 = as.matrix(fixef(fit_drg_est))[5:6] # true estimate = -1.9812346 -0.9895642
  D1 = as.matrix(VarCorr(fit_drg_est)$subj)[2:3, 2:3] 
  
  beta2 = as.matrix(fixef(fit_pbo_est))[2:3]
  gamma2 = as.matrix(fixef(fit_pbo_est))[5:6] # 1.994051 1.003545
  D2 = as.matrix(VarCorr(fit_pbo_est)$subj)[2:3, 2:3]
  
  b = purity_calculation(dat_est, beta1, beta2, gamma1, gamma2, D1, D2, p=2)
  return(list(purity = b$purity, Mu1 = b$Mu1, Mu2 = b$Mu2, beta1 = beta1, beta2 = beta2, 
              gamma1 = gamma1, gamma2 = gamma2, D1 = D1, D2 = D2, data = dat_est))
}
```


### The results

The generated data set:

```{R}
head(dat)
```

We can compare the purity calculated with different $\alpha$ values

```{R}
A = c(1,1,1) # true value
res = purity_function(A, varname = c('x1','x2','x3'), times = '', 
                      trt = '',
                      trtlevel = '',
                      subj = '',
                      outcome = 'y',
                      start = 0, data = dat)
sum(res$purity)

A = c(1,0,1)
res2 = purity_function(A, varname = c('x1','x2','x3'), times = '', 
                      trt = '',
                      trtlevel = '',
                      subj = '',
                      outcome = 'y',
                      start = 0, data = dat)
sum(res2$purity)
```


### Try to use *optim function* to find the max value

```{r fig.width=3, fig.height=5,echo=FALSE,  fig.align = "left"}

img <- readPNG("/Users/yaolanqiu/Desktop/NYU/rotation/Rotation2/Week12/1.png")
grid.raster(img)
```

### Try *genetic algoritm*

It returns a value:

```{R}
A = c(3.125,3.152,3.289)
res = purity_function(A, varname = c('x1','x2','x3'), times = '', 
                      trt = '',
                      trtlevel = '',
                      subj = '',
                      outcome = 'y',
                      start = 0, data = dat)
sum(res$purity)
```

However, the value is larger than the true value. 


# Scenario 2. Set df = 1. 

All the same with scenario 1 except the choice of $alpha$

Let $\alpha = [sin(\theta)sin(\theta), sin(\theta)cos(\theta), cos(\theta)]$, the indpendent variable will be the $\theta$. Let's set true $\theta = \frac{\pi}{3}$.

However, this time the *optim function* doesn't work well neither. 

```{r fig.width=5, fig.height=6,echo=FALSE,  fig.align = "left"}

img <- readPNG("/Users/yaolanqiu/Desktop/NYU/rotation/Rotation2/Week12/2.png")
grid.raster(img)
```



The genetic algorthm return an

* $\alpha \approx 1.01953$, $f(\alpha) \approx 338.2038$

* $\pi/3 \approx 1.0472$, $f(\pi/3) \approx 326.1763$

The purity vs $\alpha$ plot:

```{R echo = FALSE}
setwd('/Users/yaolanqiu/Desktop/NYU/rotation/Rotation2/Week12')
load('1.RData')
plot(seq(0,pi,0.01), res, type = 'l',
     main = 'Purity with theta',
     xlab = 'theta',ylab = 'purity')

points(pi/3, 326.1763, col = 'red', pch = 18, cex = 0.8)
points(seq(0,pi,0.01)[which(res == max(res))], 
       max(res), pch = 20, cex = 0.8, col = 'blue')
lines(c(pi/3, pi/3), c(0, 326.1763), lty = 3, col = 'red')
lines(c(seq(0,pi,0.01)[which(res == max(res))], 
        seq(0,pi,0.01)[which(res == max(res))]), 
      c(0, max(res)), lty = 2, col = 'blue')

legend('topright',  legend=c("True", "Estimated max"),
                          col=c("red", "blue"), lty=2:3, cex=0.8)

```

